---
title: "chlorophyll_NOAA_CoastWatch"
author: "Lia Domke"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Based on a R notebook (link: https://coastwatch.pfeg.noaa.gov/projects/r/timeseries-chl.html) found on the NOAA Coast Watch website. We're going to attempt to use the rerddapXtracto package to download satellite data from 3 different sources (and 1 compiled source) from the ERRDAP survey. 
This script is based on the code present in the r notebook 

# 1. Install required packages
```{r}
pkgTest <- function(x)
{
  if (!require(x,character.only = TRUE))
  {
    install.packages(x,dep=TRUE)
    if(!require(x,character.only = TRUE)) stop(x, " :Package not found")
  }
}

# create list of required packages
list.of.packages <- c("ncdf4","parsedate","rerddap","plotdap","sp","rerddapXtracto","lubridate", "maps","cmocean",
                       "mapdata", "maptools", "mapproj", 
                       "reshape2", "colorRamps", "plyr", "ggplot2", "gridExtra")

# create list of installed packages
pkges = installed.packages()[,"Package"]

for (pk in list.of.packages) {
  pkgTest(pk)
}

library(ncdf4)
library(parsedate)
library(rerddap)
library(plotdap)
library(sp)
library(rerddapXtracto)
library(lubridate)
library(maps)
library(cmocean)
library(mapdata)
library(mapproj)
library(reshape2)
library(colorRamps)
library(plyr)
library(ggplot2)
library(gridExtra)
```

# 2. Define geographic boundaries 

```{r}
xcoord <- c(-130, -135)
ycoord <- c(53, 57)

##Format Box Coordinates for cosmetics, to make a nice map title
ttext<-paste(paste(abs(xcoord), collapse="-"),"W, ", paste(ycoord, collapse="-"),"N")

```

# 3. SeaWiFS data
```{r}
# Use rerddap to get information about the dataset
dataInfo <- rerddap::info('erdSWchlamday')

# Display the dataset metadata
dataInfo
```

```{r}
# Extract the parameter name from the metadata in dataInfo
parameter <- dataInfo$variable$variable_name

# Set the altitude coordinate to zero
zcoord <- 0.

# Extract the beginning and ending dates of the dataset from the metadata in dataInfo
global <- dataInfo$alldata$NC_GLOBAL
tt <- global[ global$attribute_name %in% c('time_coverage_end','time_coverage_start'), "value", ]

# Populate the time vector with the time_coverage_start from dataInfo
# Use the "last" option for the ending date
tcoord <- c(tt[2],"last")

# Extract the timeseries data using rxtracto_3D
chlSeaWiFS<-rxtracto_3D(dataInfo,parameter=parameter,
                        tcoord=tcoord,
                        xcoord=xcoord,ycoord=ycoord,zcoord=zcoord)
```

```{r}
# Remove extraneous zcoord dimension for chlorophyll 
chlSeaWiFS$chlorophyll <- drop(chlSeaWiFS$chlorophyll)
```

# 4. MODIS data
```{r}
# Use rerddap to get information about the dataset
# if you encouter an error reading the nc file clear the rerrdap cache: 
# rerddap::cache_delete_all(force = TRUE)
dataInfo <- rerddap::info('erdMH1chlamday')
dataInfo
```

```{r}
# Extract the parameter name from the metadata in dataInfo
parameter <- dataInfo$variable$variable_name

#Extract the start and end times of the dataset from the metadata in dataInfo
global <- dataInfo$alldata$NC_GLOBAL

# Populate the time vector with the time_coverage_start from dataInfo
# Use the "last" option for the ending date
tt <- global[ global$attribute_name %in% c('time_coverage_end','time_coverage_start'), "value", ]
tcoord <- c(tt[2],"last")

# Run rxtracto_3D
chlMODIS<-rxtracto_3D(dataInfo,parameter=parameter,
                      tcoord=tcoord,
                      xcoord=xcoord,ycoord=ycoord)
```

# 4. VIIRS data

```{r}
# Use rerddap to get information about the dataset
# if you encouter an error reading the nc file clear the rerrdap cache: 
# rerddap::cache_delete_all(force = TRUE)
dataInfo <- rerddap::info('nesdisVHNSQchlaMonthly')
#dataInfo <- rerddap::info('erdVHNchlamday')  # alternate dataset to use
dataInfo
```

```{r}
## This extracts the parameter name from the metadata in dataInfo
parameter <- dataInfo$variable$variable_name

#Extract the start and end times of the dataset from the metadata in dataInfo
global <- dataInfo$alldata$NC_GLOBAL

# Populate the time vector with the time_coverage_start from dataInfo
# Use the "last" option for the ending date
tt <- global[ global$attribute_name %in% c('time_coverage_end','time_coverage_start'), "value", ]
tcoord <- c(tt[2],"last")
#tcoord <- c(tt[1], tt[2])

# Run rxtracto_3D
chlVIIRS<-rxtracto_3D(dataInfo,parameter=parameter,
                      tcoord=tcoord,
                      xcoord=xcoord,ycoord=ycoord,zcoord=zcoord)

## Remove extraneous zcoord dimension for chlorophyll 
chlVIIRS$chlor_a <- drop(chlVIIRS$chlor_a)
#chlVIIRS$chla <- drop(chlVIIRS$chla)
```

# 5. Create timeseries of mean monthly data
```{r}
## Spatially average all the data within the box for each dataset.
## The c(3) indicates the dimension to keep - in this case time 
chlSeaWiFS$avg <- apply(chlSeaWiFS$chlorophyll, c(3),function(x) mean(x,na.rm=TRUE))
chlMODIS$avg <- apply(chlMODIS$chlorophyll, c(3),function(x) mean(x,na.rm=TRUE))
chlVIIRS$avg <- apply(chlVIIRS$chlor_a, c(3),function(x) mean(x,na.rm=TRUE))
#chlVIIRS$avg <- apply(chlVIIRS$chla, c(3),function(x) mean(x,na.rm=TRUE))

## Temporally average all of the data into one map 
## The c(1,2) indicates the dimensions to keep - in this case latitude and longitude  
chlSeaWiFS$avgmap <- apply(chlSeaWiFS$chlorophyll,c(1,2),function(x) mean(x,na.rm=TRUE))
chlMODIS$avgmap <- apply(chlMODIS$chlorophyll,c(1,2),function(x) mean(x,na.rm=TRUE))
chlVIIRS$avgmap <- apply(chlVIIRS$chlor_a,c(1,2),function(x) mean(x,na.rm=TRUE))
#chlVIIRS$avgmap <- apply(chlVIIRS$chla,c(1,2),function(x) mean(x,na.rm=TRUE))
```

# 6. Plot time series
```{r}
## To print out a file uncomment the png command and the dev.off command
##png(file="CHL_timeseries.png", width=10,height=7.5,units="in",res=500)
plot(as.Date(chlSeaWiFS$time), chlSeaWiFS$avg, 
     type='b', bg="blue", pch=21, xlab="", cex=.7,
     xlim=as.Date(c("1997-01-01","2019-01-01")),
     ylim=c(0,13),
     ylab="Chlorophyll", main=ttext)
#axis(2)

# Now add MODIS and VIIRS  data 
points(as.Date(chlMODIS$time), chlMODIS$avg, type='b', bg="red", pch=21,cex=.7)
points(as.Date(chlVIIRS$time), chlVIIRS$avg, type='b', bg="black", pch=21,cex=.7)

text(as.Date("1997-03-01"),6, "SeaWiFS",col="blue", pos=4)
text(as.Date("1997-03-01"),4, "MODIS",col="red", pos=4)
text(as.Date("1997-03-01"),5, "VIIRS",col="black", pos=4)

#dev.off() # This closes the png file if its been written to 
```

# 7. ESA OCCI Data
```{r}
# Reading in three datasets, which  have different datset attributes (ie parameter 
# name and the presence or absence of an altitude field) is cumbersome.  ESA makes 
# a "mission-less" product, which seemlessly integrates data from all these sensors 
# into one.  So lets redo this exercise using this dateset instead and compare the results.  
rerddap::cache_delete_all(force = TRUE)
dataInfo <- rerddap::info('pmlEsaCCI31OceanColorMonthly')

# This identifies the parameter to choose - there are > 60 in this dataset1 
parameter <- 'chlor_a'

global <- dataInfo$alldata$NC_GLOBAL
tt <- global[ global$attribute_name %in% c('time_coverage_end','time_coverage_start'), "value", ]
tcoord <- c(tt[2],"last")
# if you encouter an error reading the nc file clear the rerrdap cache: 


chlOCCCI<-rxtracto_3D(dataInfo,parameter=parameter,
                      tcoord=tcoord,
                      xcoord=xcoord,ycoord=ycoord)

# Now spatially average the data into a timeseries
chlOCCCI$avg <- apply(chlOCCCI$chlor_a, c(3),function(x) mean(x,na.rm=TRUE))

# Now temporally average the data into one map 
chlOCCCI$avgmap <- apply(chlOCCCI$chlor_a,c(1,2),function(x) mean(x,na.rm=TRUE))
```

Add ESA OCCI data to the plot
```{r}
## Plot SeaWIFS
plot(as.Date(chlSeaWiFS$time), chlSeaWiFS$avg, 
     type='b', bg="blue", pch=21, xlab="", cex=.7,
     xlim=as.Date(c("1997-01-01","2019-01-01")),
     ylim=c(0, 10),
     ylab="Chlorophyll", main=ttext)
#axis(2)

## Add MODIS, VIIRS and OCCCI data 
points(as.Date(chlMODIS$time), chlMODIS$avg, type='b', bg="red", pch=21,cex=.7)
points(as.Date(chlVIIRS$time), chlVIIRS$avg, type='b', bg="black", pch=21,cex=.7)
points(as.Date(chlOCCCI$time), chlOCCCI$avg, type='b', bg="green", pch=21,cex=.5)
## Add text annotation for legend
text(as.Date("1997-03-01"),4, "SeaWiFS",col="blue", pos=4)
text(as.Date("1997-03-01"),4.5, "MODIS",col="red", pos=4)
text(as.Date("1997-03-01"),5.0, "VIIRS",col="black", pos=4)
text(as.Date("1997-03-01"),5.5, "OC-CCI",col="green", pos=4)

#dev.off() # This closes the png file if its been written to 
```

# Make maps of average chlorophyll for each satellite mission
```{r}


# Put arrays into format for ggplot
melt_map <- function(lon,lat,var) {
  dimnames(var) <-list(Longitude=lon, Latitude=lat)
  ret <- melt(var,value.name="Chl")
}

# Loop for making 4 maps
datasetnames <- c("SeaWiFS","MODIS","VIIRS","OC-CCI")
ak <- map_data("world2Hires", "USA:alaska")

plot_list = list()

for(i in 1:4) {
  
  if(i == 1) chl <- chlSeaWiFS
  if(i == 2) chl <- chlMODIS
  if(i == 3) chl <- chlVIIRS
  if(i == 4) chl <- chlOCCCI
  
   chlmap <- melt_map(chl$longitude, chl$latitude, chl$avgmap)

   p = ggplot(
     data = chlmap, 
     aes(x = Longitude, y = Latitude, fill = log10(Chl))) +
         geom_tile(na.rm=T) +
         geom_polygon(data = ak, aes(x=long, y = lat, group = group), fill = "grey80") +
         theme_bw(base_size = 12) + ylab("Latitude") + xlab("Longitude") +
         coord_fixed(1.3, xlim = xcoord, ylim = ycoord) +
         scale_fill_gradientn(colours = rev(rainbow(12)), 
                              na.value = NA, limits=c(-1,1)) +
         ggtitle(paste("Average", datasetnames[i])
      ) 

  plot_list[[i]] = p
}

# Now print out maps into a png file.  Can't use par function with **ggplpot** to get 
# multiple plots per page.  Here using a function in the **gridExtra** package

#png(file="CHL_averagemaps.png")
library(grid)
grid.arrange(plot_list[[1]],plot_list[[2]],plot_list[[3]],plot_list[[4]], nrow = 2)

# dev.off # if you'd like to write the figures to a png
```

# Make map of average chlorophy using only 3 sources
```{r}

# Put arrays into format for ggplot
melt_map <- function(lon,lat,var) {
  dimnames(var) <-list(Longitude=lon, Latitude=lat)
  ret <- melt(var,value.name="Chl")
}

# Loop for making 4 maps
datasetnames <- c("SeaWiFS","MODIS","VIIRS")
ak <- map_data("world2Hires", "USA:alaska")

plot_list = list()

for(i in 1:3) {
  
  if(i == 1) chl <- chlSeaWiFS
  if(i == 2) chl <- chlMODIS
  if(i == 3) chl <- chlVIIRS
  
   chlmap <- melt_map(chl$longitude, chl$latitude, chl$avgmap)

   p = ggplot(
     data = chlmap, 
     aes(x = Longitude, y = Latitude, fill = log10(Chl))) +
         geom_tile(na.rm=T) +
         geom_polygon(data = ak, aes(x=long, y = lat, group = group), fill = "grey80") +
         theme_bw(base_size = 12) + ylab("Latitude") + xlab("Longitude") +
         coord_fixed(1.3, xlim = xcoord, ylim = ycoord) +
         scale_fill_gradientn(colours = rev(rainbow(12)), 
                              na.value = NA, limits=c(-1,1)) +
         ggtitle(paste("Average", datasetnames[i])
      ) 

  plot_list[[i]] = p
}

# Now print out maps into a png file.  Can't use par function with **ggplpot** to get 
# multiple plots per page.  Here using a function in the **gridExtra** package

#png(file="CHL_averagemaps.png")
library(grid)
grid.arrange(plot_list[[1]],plot_list[[2]],plot_list[[3]], nrow = 2)

# dev.off # if you'd like to write the figures to a png
```


# 8. Use a different palette
The rainbow palette creates a visual bias of the data. Use a different colour palette called *cmocean*

```{r}

# Put arrays into format for ggplot
melt_map <- function(lon,lat,var) {
  dimnames(var) <-list(Longitude=lon, Latitude=lat)
  ret <- melt(var,value.name="Chl")
}

# Loop for making 4 maps
datasetnames <- c("SeaWiFS","MODIS","VIIRS","OC-CCI")
ak <- map_data("world2Hires", "USA:alaska")

plot_list = list()

for(i in 1:4) {
  
  if(i == 1) chl <- chlSeaWiFS
  if(i == 2) chl <- chlMODIS
  if(i == 3) chl <- chlVIIRS
  if(i == 4) chl <- chlOCCCI
  
   chlmap <- melt_map(chl$longitude, chl$latitude, chl$avgmap)

   p = ggplot(
     data = chlmap, 
     aes(x = Longitude, y = Latitude, fill = log10(Chl))) +
         geom_tile(na.rm=T) +
         geom_polygon(data = ak, aes(x=long, y = lat, group = group), fill = "grey80") +
         theme_bw(base_size = 12) + ylab("Latitude") + xlab("Longitude") +
         coord_fixed(1.3, xlim = xcoord, ylim = ycoord) +
         scale_fill_gradientn(colours = cmocean('algae')(9), 
                              na.value = NA, limits=c(-1,1)) +
         ggtitle(paste("Average", datasetnames[i])
      ) 

  plot_list[[i]] = p
}

# Now print out maps into a png file.  Can't use par function with **ggplpot** to get 
# multiple plots per page.  Here using a function in the **gridExtra** package

png(file="CHL_averagemaps.png")
library(grid)
grid.arrange(plot_list[[1]],plot_list[[2]],plot_list[[3]],plot_list[[4]], nrow = 2)

 dev.off # if you'd like to write the figures to a png

```

## visualize VIDRIS data only 
Runs from 2012 to present

```{r}
q <- ggplot(data= chlmap, 
            aes(x = Longitude, y = Latitude, fill = log10(Chl))) + 
  geom_tile(na.rm = T) +
  geom_polygon(data = ak, aes(x = long, y = lat, group = group), fill = "grey80") +
  theme_bw(base_size = 12) +
  ylab("Latitude") +
  xlab("Longitude") +
  coord_fixed(1.3, xlim = xcoord, ylim = ycoord) +
  scale_fill_gradientn(colours = cmocean('algae')(9),
                       na.value = NA, limits = c(-1, 1)) +
  ggtitle("Average VIIRS")
  
q
```

```{r}
## To print out a file uncomment the png command and the dev.off command
##png(file="CHL_timeseries.png", width=10,height=7.5,units="in",res=500)
chlVIIRS$date <- as.Date(chlVIIRS$time)

plot(chlVIIRS$date, chlVIIRS$avg, 
     type='b', bg="blue", pch=21, xlab="Date", cex=.7,
     ylim=c(0,7),
     ylab="Chlorophyll", main=ttext)
#axis(2)

#dev.off() # This closes the png file if its been written to 
library(tidyverse)
library(lubridate)
glimpse(chlVIIRS)

df1 <- data.frame(chlVIIRS$avg)
df1$date <- chlVIIRS$date
df1$Month_Yr <- format(as.Date(df1$date), "%Y-%m")


ggplot(data = df1, mapping = aes(x = Month_Yr, y = chlVIIRS.avg, group = 1)) +
  geom_line(stat = "identity") +
  geom_point() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text = element_text(size = 12))

```


